Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-11 eurocg/paper.tex

The idea of drawing edges only partially has been formalized in graph drawing as follows~\cite{bk-ecbe-12}. 
A \emph{partial edge drawing (PED)} is a graph drawing that maps vertices to points and edges to pairs of crossing-free edge stubs of positive length pointing towards each other.
These edge stubs are obtained by erasing one contiguous central piece of the straight-line segment connecting the two endpoints of each edge.
In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn, see Fig.~\ref{fig:examples}.
More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.

@@

A \emph{symmetric partial edge drawing} (SPED) $D$ of $\Gamma$ draws a fraction $0 < f_s \le 1$ of each edge $s = uv \in S$ by drawing two symmetric edge stubs at $u$ and $v$ of length $f_s \cdot |s|/2$ each. 
No two stubs in $D$ may intersect.
The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
In the problem \maxsped the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.

@@

			\centering
\includegraphics{Reduction_visual}
\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
\label{fig:gadget_connection}
%	\end{wrapfigure}		
		\end{figure}
	All segments in the gadgets are of length 5.
	We use pairs of intersecting segments, alternatingly colored red and green.
	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).

@@

	For each clause we construct a triple of mutually intersecting segments, see the gadget on yellow background in the upper part of Fig.~\ref{fig:gadget_connection}.
	Again, their intersection points are at distance 1 from the endpoints.
	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most 1.
	Hence the maximum amount of ink in a SPED contributed by a clause gadget is 9.

@@

	For the correctness of the reduction, let $L$ be the ink value obtained by counting 7 for each red-green segment pair and 9 for each clause gadget. 
	First assume that $\phi$ has a satisfying truth assignment and put each variable gadget in its corresponding state.
	For each clause, select exactly one literal with value \emph{true} in the satisfying truth assignment. 
	We draw the clause segment that connects to the selected literal as a full segment and the other two as length-1 stubs.
	Recall that the literal paths are oriented from the variable gadget to the clause gadget.
	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.
	Hence the first segment of the path must be a full segment.
	But because the variable is in the state that sets the literal to \emph{true}, the intersecting variable segment is drawn as two stubs and the path configuration is valid.
	For the two non-selected literals, we can draw the last segments of their paths as full segments, as well as every segment at an even position, while the segments at odd positions are drawn as stubs. 
	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value~$L$.

	Conversely, assume that we have a SPED with ink value $L$. 
	By construction, every red-green segment pair and every clause gadget must contribute its respective maximum ink value.
	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
	We claim that this truth assignment satisfies $\phi$.
	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments. 
	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
	But then by construction the selected variable value sets our literal to \emph{true}.	
\end{proof}

@@

Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:

@@

The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.

>>>>

The idea of drawing edges only partially has been formalized in graph drawing as follows~\cite{bk-ecbe-12}. 
A \emph{partial edge drawing (PED)} is a graph drawing that maps vertices to points and edges to pairs of crossing-free edge stubs of positive length pointing towards each other.
These edge stubs are obtained by erasing one contiguous central piece of the straight-line segment connecting the two endpoints of each edge.
In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn (see Fig.~\ref{fig:examples}).
More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.

@@

A \emph{symmetric partial edge drawing} (SPED) $D$ of $\Gamma$ draws a fraction $0 < f_s \le 1$ of each edge $s = uv \in S$ by drawing two symmetric edge stubs at $u$ and $v$ of length $f_s \cdot |s|/2$ each. 
No two stubs in $D$ may intersect.
The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
In the problem \maxsped, the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.

@@

			\centering
\includegraphics{Reduction_visual}
\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
\label{fig:gadget_connection}
%	\end{wrapfigure}		
		\end{figure}
	All segments in the gadgets are of length 5.
	We use pairs of intersecting segments, alternatingly colored red and green.
	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
	Thus, the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).

@@

	For each clause we construct a triple of mutually intersecting segments, see the gadget on yellow background in the upper part of Fig.~\ref{fig:gadget_connection}.
	Again, their intersection points are at distance~1 from the endpoints.
	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most~1.
	Hence, the maximum amount of ink in a SPED contributed by a clause gadget is~9.

@@

	For the correctness of the reduction, let $L$ be the ink value obtained by counting 7 for each red-green segment pair and 9 for each clause gadget. 
	First assume that $\phi$ has a satisfying truth assignment and put each variable gadget in its corresponding state.
	For each clause, select exactly one literal with value \emph{true} in the satisfying truth assignment. 
	We draw the clause segment that connects to the selected literal as a full segment and the other two as length-1 stubs.
	Recall that the literal paths are oriented from the variable gadget to the clause gadget.
	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.
	Hence, the first segment of the path must be a full segment.
	But because the variable is in the state that sets the literal to \emph{true}, the intersecting variable segment is drawn as two stubs and the path configuration is valid.
	For the two non-selected literals, we can draw the last segments of their paths as full segments, as well as every segment at an even position, while the segments at odd positions are drawn as stubs. 
	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value~$L$.

	Conversely, assume that we have a SPED with ink value $L$. 
	By construction, every red-green segment pair and every clause gadget must contribute its respective maximum ink value.
	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
	We claim that this truth assignment satisfies $\phi$.
	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments. 
	In particular, the first segment of the literal path is fully drawn and therefore, in a SPED with ink $L$, the intersecting segment of the variable cannot be fully drawn.
	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
	But then by construction, the selected variable value sets our literal to \emph{true}.	
\end{proof}

@@

Further, we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:

@@

The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$, so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
